---
title: "graphs student loans"
author: "Camille Bergeron"
date: "2022-09-12"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

# Student Loans

## United States

+ the numbers following the graphs count the number of blue dots (i.e. the geogrpahies where the communities of color have a higher median or share than majority white communities)
+ the fraction after is the proportion to all of the available data

```{r}
# Student loan script
# this is to do exploratory data anlyssi for student loan stuff

# Camille Bergeron
# September 2022

rm(list = ls())
library(tidyverse)
library(plotly)
library(readxl)

dat_raw <- read_xlsx("C:/Users/camilleb/Box/Research/Active Projects/Higher Education in Boston_2022/03. Data/county_dia_student_ 7 Jun 2022.xlsx")

# batch making the columns numeric 

dat <- sapply(dat_raw %>% select(-c(NAME, state_name)), as.numeric) %>%
  as_tibble()

dat$NAME = dat_raw$NAME
dat$state_name = dat_raw$state_name

# graphs

dat %>%
  # filter(state_name %in% state.name[state.region == "Northeast"]) %>%
  mutate(highlight = ifelse(`Median Student Loan Debt, Communties of Color`>`Median Student Loan Debt, Majority White Communities`, 1, 0)) %>%
  ggplot(aes(x = `Median Student Loan Debt, Communties of Color`, y = `Median Student Loan Debt, Majority White Communities`)) +
  geom_point(aes(group = state_name, color = highlight)) +
  geom_abline(intercept = 0, slope = 1)

n_median <- sum(dat %>% mutate(highlight = ifelse(`Median Student Loan Debt, Communties of Color`>`Median Student Loan Debt, Majority White Communities`, 1, 0)) %>% select(highlight), na.rm = TRUE)
n_median_tot <- sum(dat %>% mutate(highlight = ifelse(`Median Student Loan Debt, Communties of Color`>`Median Student Loan Debt, Majority White Communities`, 1, 1)) %>% select(highlight), na.rm = TRUE)

n_median
n_median / n_median_tot

dat %>%
  # filter(state_name %in% state.name[state.region == "Northeast"]) %>%
  mutate(highlight = ifelse(`Share with Student Loan Debt, Communities of Color`>`Share with Student Loan Debt, Majority White Communities`, 1, 0)) %>%
  ggplot(aes(x = `Share with Student Loan Debt, Communities of Color`, y = `Share with Student Loan Debt, Majority White Communities`)) +
  geom_point(aes(group = state_name, color = highlight)) +
  geom_abline(intercept = 0, slope = 1)

n_share <- sum(dat %>% mutate(highlight = ifelse(`Share with Student Loan Debt, Communities of Color`>`Share with Student Loan Debt, Majority White Communities`, 1, 0)) %>% select(highlight), na.rm = TRUE)
n_share_tot <- sum(dat %>% mutate(highlight = ifelse(`Share with Student Loan Debt, Communities of Color`>`Share with Student Loan Debt, Majority White Communities`, 1, 1)) %>% select(highlight), na.rm = TRUE)

n_share
n_share/n_share_tot


```

# Massachusetts

+ Suffolk county is highlighted in blue

```{r}
dat %>%
  filter(state_name == "Massachusetts") %>%
  mutate(highlight = ifelse(NAME == "Suffolk County, Massachusetts", 1, 0)) %>%
  ggplot() +
  geom_point(aes(y = `Median Student Loan Debt, All`, x = `Average household income, 2019 (ACS)`, color = highlight))+
  theme(legend.position = "none") 


```

+ the outlier on the bottom-right is nantucket county 

What's interetsing about this is the measurement of income and not wealth  

# Race

```{r}

# dat %>%
#  filter(state_name == "Massachusetts") %>%
#  mutate(median_income_2019_acs = c(85042, 104007, 116351, 51952, 58895, ))



```

## Dumbbell graph

```{r}
dat %>%
  filter(state_name == "Massachusetts") %>% 
  mutate(NAME = sub(",.*", "", NAME)) %>%
  mutate(highlight = ifelse(`Median Student Loan Debt, Communties of Color`>`Median Student Loan Debt, Majority White Communities`, "Higher", "Lower")) %>%
  ggplot() +
  geom_segment(aes(x = `Median Student Loan Debt, Communties of Color`, y = NAME, 
                   xend = `Median Student Loan Debt, Majority White Communities`, yend = NAME,
                   color = highlight), 
               size = 3, 
               alpha = 0.25) +
  scale_color_manual(values = c("#003450", "#0fa6b4")) +
  geom_point(aes(x = `Median Student Loan Debt, Communties of Color`, y = NAME), 
             color = "#0fa6b4", size = 3)+
  geom_point(aes(x = `Median Student Loan Debt, Majority White Communities`, y = NAME), 
             color = "#003450", size = 3) +
  theme(legend.position = "none", 
        title = element_text(colour = "#003450"), 
        axis.text = element_text(colour = "#003450")) +
  labs(title = "Student Loan Debt", x = "Median Student Loan Debt ($)", y = "")

  
```

